Can Auxiliary Indicators Improve COVID-19 Forecasting and Hotspot Prediction?

### 2021-06-17 forecast/paper/paper.tex
- Talk about Google latency - Add interesting vintage vs. finalized insights - Fix typos

In our companion paper, we analyze correlations between various indicators and   
COVID case rates. These correlations are natural summaries of the
contemporaneous association between an indicator and COVID cases, but they fall 
short of delivering a satisfactory answer to the question that motivates the
current article: is the information contained in an indicator demonstrably
useful for the prediction tasks we care about? For such a question, lagged
correlations (e.g., measuring the correlation between an indicator and COVID
case rates several days in the future) move us in the right direction, but still
do not move us all the way there. The question about \textit{utility for 
  prediction} is focused on a much higher standard than simply asking 
about correlations; to be useful in forecast or hotspot models, an indicator
must provide relevant information that is not otherwise contained in past values
of the case rate series itself. We will assess this in the most direct way
possible: by inpsecting the difference in predictive performance of simple
autoregressive models trained with and without access to past values of a
particular indicator.

@@

Lastly, our treatment of the Google-AA signal is different from the rest.
Because Google's team did not start publishing this signal until early
September, 2020, we do not have true vintage data before then.  However, this
signal is never revised after initial publication (confirmed via personal
communication with the Google team that produces this signal), so we simply use
finalized values for it in our analysis.

@@

Over the prediction period, all auxiliary indicators are available (in the
proper vintage sense) for all locations and prediction times, except for the
Google-AA signal, which is only observed for an average of 105 (of 306) HRRs.
This is due to the fact that that the COVID-19 search trends data is released
after applying differential privacy techniques \cite{Bavadekar:2020}, and a
missing signal value means that the level of noise added in the differential
privacy mechanism is high compared to the underyling search count.  In other
words, values of the Google-AA signal are clearly \textit{not} missing at
random.  It seems most appropriate to impute missing values by zero, and this 
is what we do in our analysis.

@@

Finally, to quantify the importance of training and making predictions using
proper vintage data, we ran a parallel set of forecasting and hotspot prediction 
experiments using finalized data. The results, given in the supplement, show
that training and making predictions on finalized data can result in overly
optimistic estimates of true test-time performance (up to 10\% better in terms
of average WIS or AUC).
% TODO note what we said here about the supplement

Code to reproduce all results (which uses the \texttt{evalcast} R package) can 
be found at
\url{https://www.github.com/cmu-delphi/covidcast-pnas/forecast/code/}.

@@

One might ask if the benefits we observe in forecasting and hotspot have
anything to do with the actual indicator themselves; could it instead be that
the indicators are simply providing ``implicit regularization'' on top of the
basic AR model, when we include them as lagged features in
\eqref{eq:forecast_ar_x}, \eqref{eq:hotspot_ar_x}?

@@

Can axuiliary indicators improve COVID-19 forecasting and hotspot prediction
models?  Our answer, based on analyzing five auxiliary indicators from the  
COVIDcast API (defined using from medical insurance claims, internet-based
surveys, and Google search trends) is undoubtedbly ``yes''. However, there are 
levels of nuance to such an answer that must be explained.  None of the
indicators that we have investigated appear to be the ``silver bullet'' that 
one  might have hoped for, revolutionizing the tractability of the prediction 
problem, rendering it easy when it was once hard (in the absence of auxiliary 
information).  Rather, the gains in accuracy offered by the indicator models 
(over an autoregressive model based on past case rates along) appear to be
nontrivial, and consistent across modes of analysis, but modest.  In
forecasting, the indicator models appear to be most useful in periods in which 
case rates are flat or trending down, rather than periods in which case rates
are trending up (as one might hope to see, as the hypothetical benefit provided
by an idealized ``leading indicator'').

>>>>

In our companion paper, we analyze correlations between various indicators and   
COVID case rates. These correlations are natural summaries of the
contemporaneous association between an indicator and COVID cases, but they fall 
short of delivering a satisfactory answer to the question that motivates the
current article: is the information contained in an indicator demonstrably
useful for the prediction tasks we care about? For such a question, lagged
correlations (e.g., measuring the correlation between an indicator and COVID
case rates several days in the future) move us in the right direction, but still
do not move us all the way there. The question about \textit{utility for 
  prediction} is focused on a much higher standard than simply asking 
about correlations; to be useful in forecast or hotspot models, an indicator
must provide relevant information that is not otherwise contained in past values
of the case rate series itself. We will assess this in the most direct way
possible: by inspecting the difference in predictive performance of simple
autoregressive models trained with and without access to past values of a
particular indicator.

@@

Lastly, our treatment of the Google-AA signal is different from the rest.
Because Google's team did not start publishing this signal until early
September, 2020, we do not have true vintage data before then.
Furthermore, the latency of the signal was always at least one week
through 2020.  However, this
signal is never revised after initial publication (confirmed via personal
communication with the Google team that produces this signal) and
furthermore the latency of the signal is not an unavoidable property
of the data type, so we simply use finalized values for it in our analysis.

@@

Over the prediction period, all auxiliary indicators are available (in the
proper vintage sense) for all locations and prediction times, except for the
Google-AA signal, which is only observed for an average of 105 (of 306) HRRs.
This is due to the fact that the COVID-19 search trends data is released
after applying differential privacy techniques \cite{Bavadekar:2020}, and a
missing signal value means that the level of noise added in the differential
privacy mechanism is high compared to the underyling search count.  In other
words, values of the Google-AA signal are clearly \textit{not} missing at
random.  It seems most appropriate to impute missing values by zero, and this 
is what we do in our analysis.

@@

Finally, to quantify the importance of training and making predictions using
proper vintage data, we ran a parallel set of forecasting and hotspot prediction 
experiments using finalized data. The results, given in the supplement, show
that training and making predictions on finalized data can result in overly
optimistic estimates of true test-time performance (up to 10\% better in terms
of average WIS or AUC). Furthermore, since indicators can have
greatly different backfill profiles, the use of finalized data in
retrospective evaluations can even change the relative ordering of
methods.  For example, CHNG-CLI and DV-CLI
when trained on finalized data perform very similarly for
the forecasting task.  This makes sense since they are both claims-based indicators measuring supposedly the
same thing.  However, DV-CLI outperforms CHNG-CLI on vintage data,
reflecting its less severe backfill profile.

% TODO note what we said here about the supplement

Code to reproduce all results (which uses the \texttt{evalcast} R package) can 
be found at
\url{https://github.com/cmu-delphi/covidcast-pnas/tree/main/forecast/code}.

@@

One might ask if the benefits we observe in forecasting and hotspot
prediction have
anything to do with the actual indicator themselves; could it instead be that
the indicators are simply providing ``implicit regularization'' on top of the
basic AR model, when we include them as lagged features in
\eqref{eq:forecast_ar_x}, \eqref{eq:hotspot_ar_x}?

@@

Can auxiliary indicators improve COVID-19 forecasting and hotspot prediction
models?  Our answer, based on analyzing five auxiliary indicators from the  
COVIDcast API (defined using from medical insurance claims, internet-based
surveys, and Google search trends) is undoubtedly ``yes''. However, there are 
levels of nuance to such an answer that must be explained.  None of the
indicators that we have investigated appear to be the ``silver bullet'' that 
one  might have hoped for, revolutionizing the tractability of the prediction 
problem, rendering it easy when it was once hard (in the absence of auxiliary 
information).  Rather, the gains in accuracy offered by the indicator models 
(over an autoregressive model based on past case rates along) appear to be
nontrivial, and consistent across modes of analysis, but modest.  In
forecasting, the indicator models appear to be most useful in periods in which 
case rates are flat or trending down, rather than periods in which case rates
are trending up (as one might hope to see, as the hypothetical benefit provided
by an idealized ``leading indicator'').
